---
title: "Whois Demo"
output: 
    html_document:
        keep_md: true
---

We will first install a [whois](https://en.wikipedia.org/wiki/WHOIS) utility 
if we are running Windows, as this operating system does not come with this 
tool by default. Other popular operating systems like OSX and Linux come with
`whois` pre-installed, so we will not perform this step on those systems.

```{r}
if (Sys.info()["sysname"] == "Windows") {
    # Get the whois.exe file for Windows from SysInternals.
    exefile <- ".\\whois.exe"   # This syntax will work in DOS and Bash.
    
    # Only download and extract the utility if we don't already have it.
    if (! file.exists(exefile) == TRUE) {
        zipfile <- "WhoIs.zip"
        if (! file.exists(zipfile) == TRUE) {
            url <- "http://download.sysinternals.com/files/WhoIs.zip"
            download.file(url, zipfile)
            
            # Run this utility once (first time) to agree to the EULA.
            system(paste(exefile, "--help"), intern = FALSE)
        }
        unzip(zipfile)
    }
} else {
    exefile <- "whois"
}
```

Run the `whois` command with the domain `who.int` as a test.

```{r}
# Try out the whois utility, removing any carriage-return (\r) characters.
whois.data <- gsub(pattern = "\\r", replacement = "", 
                   x = system(paste(exefile, "who.int"), intern = TRUE))

# Take a quick look at the results.
length(whois.data)
head(whois.data, 10)
```

Read the data into a `data.frame`.

```{r}
# clean up the the whois data and read into a data.frame.

# Use the vertical bar symbol (|) as the delimiter instead of the colon (:).
whois.data <- gsub(pattern = ":[[:space:]]", replacement = "|", x = whois.data)

# Remove lines which do not contain the delimiter.
whois.data <- whois.data[grepl(x = whois.data, pattern = "\\|")]

# Take a quick look at the results.
length(whois.data)
head(whois.data, 10)

# Take a look at the last few lines.
tail(whois.data, 10)

# Remove any '>>>' and '<<<' strings, if present.
whois.data <- gsub("[[:space:]]?[<>]{3}[[:space:]]?", "", whois.data)

# Check the last line again.
whois.data[length(whois.data)]

# Read the file into a data.frame (df).
whois.df <- read.table(text = whois.data, sep = "|", as.is = TRUE)
names(whois.df) <- c("variable", "value")
```

Load the `knitr` package so that we can use the `kable()` function.

```{r, message=FALSE}
library(knitr)
```

Produce a formatted table from the `data.frame` using `kable()`.

```{r, results='asis'}
kable(whois.df)
```